### 2024-06-16 manuscript.Rmd
update

---
title: "Performance of Individual Audit Filters in Predicting Opportunities for Improvement
  in Adult Trauma Patients"
subtitle: A single-center registry-based retrospective cohort
output: html_document
author: "Josef Al-Khalili"
bibliography: bibliography.bib
csl: vancouver.csl
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

## Background

Trauma is responsible for 9% of global deaths annually
and has a significant impact on individuals and society. Quality
Improvement (QI) programs seek to minimize these adverse outcomes using
Audit Filters (AFs) and interdisciplinary Morbidity and Mortality
Conferences (MMC). AFs are based on criteria for standard care and can
function as a screening tool of eligible patient cases for discussion at
MMCs. One of the goals with MMCs is to determine potential Opportunities
for Improvement (OFI).This study aimed to evaluate the
performance of individual AFs in predicting OFIs and to assess their
relevance as a screening tool for Morbidity and Mortality Conferences
(MMC).

## Methods

Register data from 8,309 patient cases
were collected for analysis. The sensitivity, specificity and Area Under
the Receiver Operating Characteristics Curve (AUC) was calculated to
assess performance. An empirical bootstrapping test of AUC values was
conducted to compare the performance of each AF to coincidence (AUC =
0.5) and minimum AUC for diagnostic tests (AUC = 0.8).

## Results

Four
out of the 11 AFs assessed had an AUC of 0.5 (p \> 0.05). The remaining
AFs demonstrated a correlation with OFI, with AUC ranging from 0.51 to
0.63. The AF "\> 60 min until first intervention" exhibited the highest
AUC (0.63, 95% CI [0.61-0.65]).

## Conclusion

No individual AF seemed to
perform well in predicting OFI, which suggest rethinking the current
screening method for MMCs.

# Introduction

Trauma, defined as a physical injury and the body's associated response,
causes 8-9% of global deaths annually [@world2014injuries;
@roth2018global], disproportionately affecting younger individuals in low- and
middle-income countries, primarily due to road traffic injuries
[@vos2020global; @roth2018global; @SweTrau2022Arsrapport]. Trauma also ranks highest in disability-adjusted life-years (DALYs) [@vos2020global] and increases the risk of mental disorders and cognitive decline [@kessler1995posttraumatic;
@north1999psychological; @muscatelli2017prevalence; @newnam2014impacts;
@visser2017course], impacting individuals, communities and
society at large [@newnam2014impacts]. Road traffic injuries alone cost approximately 2% of gross domestic product
(GDP) in high-income countries [@world2014injuries].

To reduce trauma burden, quality improvement (QI) programs have been
established by the World Health Organization (WHO) and the International
Association for Trauma Surgery and Intensive Care (IATSIC)
[@world2009guidelines]. These programs form the core of current trauma
care systems and their development around the world
[@American2022resources]. QI programs consist of a series of recommended
techniques aiming to facilitate discussion and analysis of the trauma
process in order to take corrective action for future success. The
following techniques are commonly used: a multidisciplinary mortality
and morbidity conference (MMC), a preventable death review panel, and
the use of AFs also known as quality indicators [@world2009guidelines].

Trauma care is a multidisciplinary, complex and time-critical health
care service provided by hospitals, often trauma centres.
[@American2022resources]. Trauma centres have a significantly lower
mortality rate compared to non-trauma centres, and the maturity of a
trauma centre correlates with lower mortality, demonstrating the
importance of high-quality, specialised trauma care
[@alharbi2021effectiveness; @mann1999systematic;
@candefjord2020mortality]. Despite this, even established trauma care
systems are prone to errors due to the complex nature of trauma and its
susceptibility to delays [@nikouline2021errors; @vioque2014classifying;
@o2013opportunities]. Studies between 1990 and 2014 have shown a pooled
preventable death rate of 20%, with more recent studies showing a lower
preventable death rate [@kwon2014pooled]. For example, a recent study in
a Swedish Level I trauma centre found a preventable death rate of 4%
[@ghorbani2018analysis].

### Morbidity and Mortality Conferences

The MMC is a central QI technique. It is a proven framework with the
critical components of anonymity, focusing on specific adverse events,
recognising flawed approaches, critical analysis, improvement proposals
and implementing such [@gregor2016morbidity].

The conference participants are healthcare professionals involved in
trauma care and its processes, but not necessarily the specific patient
case. The selection of patient cases varies, but common themes for
discussion are unexpected mortality, unexpected morbidity and errors
[@orlander2003morbidity].

Implemented effectively, MMCs serve as an important tool for error
reduction without being perceived as negative feedback for individual
care providers, but rather as an incentive for reporting. These
conferences promote a more comprehensive, systems-based approach that
emphasises task analysis, teamwork and quality improvement
[@lazzara2022morbidity]. The introduction of MMC has been shown to
increase morbidity and mortality reporting [@kong2016analysis].

### Audit Filters and Opportunities for Improvement

AFs, also known as quality indicators are established criteria based on
standardised care. Any deviation from the AFs should raise the attention
of health care professionals with the purpose of improving the quality
of care [@American2022resources]. An AF can involve outcomes such as
"death" or criteria related to guidelines and gold standards such as
"the placement of two large bore intravenous lines within 15 minutes
from arrival to a health care facility" [@world2009guidelines].

Previous studies have shown a reduction in trauma-related mortality
after the implementation of AFs [@chadbunchachai2001comparative;
@chadbunchachai2003study]. Per contra, newer studies seem to differ. In
2009, an extensive review by Evans et al. [@evans2009audit] sought to
determine if AFs could be used in improving processes in trauma care.
They found no studies successfully providing evidence for the use of
AFs. More than 1,500 AFs have been identified in a review by Stelfox et
al. [@stelfox2010quality]. Notably, they did not find a common set of
clearly defined, evidence-based, broadly accepted AFs for evaluating the
quality of trauma care. The perceived usefulness of audit filters still
remains high [@berg2022perceived]

Opportunities for improvement (OFIs) are the endpoint of patient care.
Following the review of individual patient cases there is a decision
made regarding the existence of an OFI [@American2022resources]. It has
been shown that such review processes are associated with high-quality
trauma care [@santana2014development]. OFIs are typically associated
with failures in initial care [@vioque2014classifying], specifically in
airway management, fluid resuscitation, haemorrhage control and chest
injury management [@sanddal2011analysis; @roy2017learning;
@o2013opportunities].

There is lacking evidence for the use of AFs for improving trauma care
[@evans2009audit; @stelfox2010quality] and the focus has been on
mortality as outcome measure [@chadbunchachai2001comparative;
@chadbunchachai2003study]. Furthermore, the use of AFs has also been
associated with high frequencies of false positive OFI predictions,
ranging from 24% to 80% [@sanddal2011analysis; @roy2017learning], and
some AFs do not appear to correlate with OFIs at all [@lewis2017delay].

```{=html}
<!-- Explain the scientific background and rationale for the
investigation being reported. End by stating the aim or specific
objectives, including any prespecified hypotheses. The introduction
should not be longer than 1-1.5 pages, typically 4-5 paragraphs, each
6-8 lines long. Support every statement with a citation -->
```
# Methods

## Study design

<!-- Present key elements of study design -->

## Setting
Karolinska University Hospital, a Trauma level I
hospital, treats around 1,500 trauma patients annually. If a patient case results in team activation or the
patient has an Injury Severity Score (ISS) of less than 9
retrospectively, it is added to the Karolinska
trauma registry which reports to SweTrau. The registry includes data on vital
signs, times, injuries, and interventions as well as patient
demographics according to the European Consensus Statement, the Utstein
template [@ringdal2008utstein]

The Karolinska trauma registry also contains a local care quality
database with specific AFs and OFIs that are determined in MMC
conferences through consensus decisions. OFIs are assessed in multiple stages,
characterised by escalating levels of scrutiny. Notably, instances of
mortality are directly referred to the multidisciplinary conference,
where, in addition to assessing OFIs, a decision is made regarding
whether the death was preventable or potentially preventable, a
classification also falling under the purview of OFIs.

From 2013 to 2017, efforts focused on identifying adverse outcomes unrelated to mortality. The review process was refined and formalized, with trauma cases initially assessed by a specialized trauma nurse for potential OFIs. In 2017, this procedure became standard practice, incorporating preliminary evaluations by specialized trauma nurses. Trauma quality data screened for AFs and cases flagged for possible care failures underwent secondary review by two specialized nurses. Potential OFIs identified in this review prompted a comprehensive evaluation of the trauma patient's case.
```{=html}
<!-- Describe the setting, locations, and relevant dates, including
periods of recruitment, exposure, follow-up, and data collection -->
```
## Participants

The study included 8,309 patients treated at Karolinska University Hospital from 2012 to 2022 who underwent OFI screening. Patients under 15 and those not screened for OFIs were excluded due to differences in clinical management for minors.

```{=html}
<!-- Cohort study: Give the eligibility criteria, and the sources and
methods of selection of participants. Describe methods of
follow-up. For matched studies, give matching criteria and number of
exposed and unexposed -->
```
```{=html}
<!-- Case-control study: Give the eligibility criteria, and the
sources and + methods of case ascertainment and control
selection. Give the rationale for the choice of cases and
controls. For matched studies, give matching criteria and the number
of controls per case -->
```
```{=html}
<!-- Cross-sectional study: Give the eligibility criteria, and the
sources and methods of selection of participants -->
```
## Variables and data sources/measurements
The outcome variable in this study is an OFI, as established by the MMCs
and furthermore treated as a dichotomous variable with "Yes - At least
one OFI identified" and "No - No OFI identified".
The 11 AFs used at Karolinska University Hospital served as the exposure variable, as shown in Table 1. Each AF is labeled "original" or "manually created." Manually created AFs were calculated using trauma registry data, while original AFs were recorded in the quality database based on criteria beyond the available registry data.

The data spanning 2012 to 2022 was retrieved from the
Karolinska University Hospital trauma registry and trauma care quality
database. Patient data on vital signs, care processes and interventions, level of
care and time aspects was retrieved from SweTrau while both exposures
(AFs) and outcome (OFI) were retrieved from the Karolinska University
Hospital trauma care quality database. All data was anonymised to protect patient privacy.
```{=html}
<!-- Clearly define all outcomes, exposures, predictors, potential
confounders, and effect modifiers. Give diagnostic criteria, if
applicable. For each variable of interest, give sources of data and
details of methods of assessment (measurement). Describe comparability
of assessment methods if there is more than one group -->
```
## Bias
Consensus decisions at MMC conferences may introduce misclassification bias. Additionally, AF flagging by specialized nurses is a manual process, potentially leading to selection bias. Healthcare professionals attending MMCs may not accurately recall key details impacting OFI decisions, introducing recall bias. Furthermore, professionals not present for specific cases might over- or underestimate OFI significance.
<!-- Describe any efforts to address potential sources of bias -->

## Study size

<!-- Explain how the study size was arrived at -->

## Quantitative variables

```{=html}
<!-- Explain how quantitative variables were handled in the
analyses. If applicable, describe which groupings were chosen and why
-->
```
## Statistical methods
The study utilized R [@R2022language] for statistical analysis. Patient characteristics between cases with and without OFI were compared using the Wilcoxon signed-rank test for continuous variables and the Chi-squared test for categorical variables.

The predictive performance of AFs was assessed by calculating sensitivity and specificity for each AF in predicting OFI. Receiver operating characteristic (ROC) curves were constructed for binary classification, and the area under the curve (AUC) was computed to compare AF performance. [@robin2011proc].

Bootstrap resampling (1,000 iterations) was applied to estimate 95% confidence intervals for sensitivity, specificity, and AUC. The significance of each AF's AUC was evaluated against chance (AUC = 0.5) and the minimum standard for diagnostic tests (AUC = 0.8) [@nahm2022receiver] using an empirical bootstrap test (1,000 iterations).

Missing values for AFs were treated as negative (i.e., "not flagged").
```{=html}
<!--

(a) Describe all statistical methods, including those used to control
for confounding

(b) Describe any methods used to examine subgroups and interactions

(c) Explain how missing data were addressed

(d)

Cohort study: If applicable, explain how loss to follow-up was addressed

Case-control study:If applicable, explain how matching of cases and
controls was addressed

Cross-sectional study: If applicable, describe analytical methods
taking account of sampling strategy

(e) Describe any sensitivity analyses
-->
```
# Results

##Participants After excluding patients under 15 years old and patients
not screened for OFIs, a total of 8,309 individuals were included in the
study. 7,797 out of 8,309 (93.8%) patient cases did not have an OFI and
512 out of 8,309 (6.2%) had an OFI. Figure 1 presents a flowchart
showcasing the exclusion and inclusion of the patient cases from the
trauma registry.

Most trauma patients were men (n = 5,755, 69%). Patient characteristics
varied depending on whether patients had an OFI or not. Respectively,
the mean age was significantly higher\
`r inline_text(tableOne, variable = pt_age_yrs, column = stat_2)` vs
`r inline_text(tableOne, variable = pt_age_yrs, column = stat_1)`,
`r inline_text(tableOne, variable = pt_age_yrs, column = "p.value")` and
the ISS was significantly higher
`r inline_text(tableOne, variable = ISS, column = stat_2)` vs
`r inline_text(tableOne, variable = ISS, column = stat_1)`,
`r inline_text(tableOne, variable = ISS, column = "p.value")`. The
number of intubations in the emergency department was significantly
higher
`r inline_text(tableOne, variable = ed_intubated, level = "Yes", column = stat_2)`
vs
`r inline_text(tableOne, variable = ed_intubated, level = "No", column = stat_1)`,
`r inline_text(tableOne, variable = ed_intubated, level = "No", column = "p.value")`.
Patients with OFI had longer times to definitive treatment from hospital
arrival compared to patients without OFI
`r inline_text(tableOne, variable = dt_ed_emerg_proc, column = stat_2)`
vs
`r inline_text(tableOne, variable = dt_ed_emerg_proc, column = stat_1)`,
`r inline_text(tableOne, variable = dt_ed_emerg_proc, column = "p.value")`.
The time to the first CT was also longer in patients with OFIs compared
to patients without
`r inline_text(tableOne, variable = dt_ed_first_ct, column = stat_2)` vs
`r inline_text(tableOne, variable = dt_ed_first_ct, column = stat_1)`,
`r inline_text(tableOne, variable = dt_ed_first_ct, column = "p.value")`.

The number of OFIs were highest in patients admitted to the intensive
care unit
`r inline_text(tableOne, variable = host_care_level, level = "Intensive care unit", column = stat_2)`
and surgical ward
`r inline_text(tableOne, variable = host_care_level, level = "Surgical ward", column = stat_2)`.

Gender `r inline_text(tableOne, variable = Gender, column = "p.value")`,
the mean systolic blood pressure
`r inline_text(tableOne, variable = ed_sbp_value, column = "p.value")`,
death at 30 days
`r inline_text(tableOne, variable = res_survival, column = "p.value")`
and trauma team activation
`r inline_text(tableOne, variable = Tr_Nivå, column = "p.value")` was
not significantly correlated with OFI.

Notably, there was a statistical significant correlation between GCS at
ED and OFI (p \< 0.001).

Table 2 shows the missing data for each AF. As mentioned in the method
section, these values were interpreted as "not flagged" with an AF. The
variables with the highest amount of missing data were "time to
definitive treatment" (n = 5,990, 72.1%) and "time to first CT" (n =
1,012, 12.2%). The smallest amount of missing data was found in
"SBP\<90" (n = 13, 0.2%).

## Performance of Individual Audit Filters

The performance of each AF was determined through a calculation of
sensitivity and specificity which is demonstrated in Table 4.

The number of patient cases flagged by each individual AF varied between
107 for "GCS\<9 and not intubated" and 3907 for "\>30 min until CT". The
AF with the highest sensitivity was "\>30 min until CT" (59.2%, 95% CI
[55.2-63.8]). The lowest sensitivity was seen in "GCS\<9 and not
intubated" (1.2%, 95% CI [0.2-2]). The AF with the highest specificity
was "GCS\<9 and not intubated" (98.7%, 95% CI [98.5-98.9]). The lowest
specificity was seen in "\>30 min until CT" (53.8%, 95% CI [52.7-54.8]).
Notably, the confidence intervals varied between sensitivity and
specificity, with sensitivity displaying a greater interval.

Table 5 summarises the performance of each individual AF in identifying
OFIs in trauma care by AUC. Four of the 11 AFs, "SBP\<90", "dead at 30
days", "GCS\<9 and not intubated" and "CPR and thoracotomy" showed a
performance in predicting OFIs equal to that of coincidence (AUC = 0.5).
The highest AUC was seen in "\>60 min until first intervention" (0.63,
95% CI [0.61-0.65]). All AFs showed significantly lower AUC compared to
the minimum standard for diagnostic tests (AUC = 0.8) (p \< 0.001).

Table 2: Comparison of missing values for each AF. Missing values is
shown as a number and percentage of the total number of patient cases
included in the study (n = 8309).\
`r tableTwo` Table 3: Patient characteristics of individuals with and
without an OFI. Missing values are shown only when they exist.
`r tableOne` Table 4: Comparison of performance for each AF in
predicting OFIs. Performance is defined by sensitivity (%) and
specificity (%). A 95% confidence interval is shown together with each
performance indicator. The number of cases flagged by each filter is
displayed.\
`r tableThree` Table 5: Comparison of performance for each AF in
predicting OFIs. Performance is defined by the area under the receiver
operating characteristics curve. P-values show the significance of AUC
compared to coincidence (AUC = 0.5) and to the minimum standard for
diagnostic tests (AUC = 0.8) [@nahm2022receiver]. The p-values were
calculated using an empirical bootstrap test. Furthermore, a 95%
confidence interval was used. `r tableFour`

```{r fig1, echo=FALSE, fig.align="center", warning=FALSE, message=FALSE, error=FALSE}
knitr::include_graphics("ofi-flowchart.pdf")
```

Figure 1: Flowchart of the patient selection process. Shows the patients
excluded and included in this study.

<!--

Participants
------------

(a) Report numbers of individuals at each stage of study—eg numbers
potentially eligible, examined for eligibility, confirmed eligible,
included in the study, completing follow-up, and analysed

(b) Give reasons for non-participation at each stage

(c) Consider use of a flow diagram

Descriptive data
----------------

(a) Give characteristics of study participants (eg demographic,
clinical, social) and information on exposures and potential
confounders

(b) Indicate number of participants with missing data for each
variable of interest

(c) Cohort study — Summarise follow-up time (eg, average and total
amount)

Outcome data
------------

Cohort study — Report numbers of outcome events or summary measures
over time

Case-control study — Report numbers in each exposure category, or
summary measures of exposure

Cross-sectional study — Report numbers of outcome events or summary
measures

Main results
------------

(a) Give unadjusted estimates and, if applicable, confounder-adjusted
estimates and their precision (eg, 95% confidence interval). Make
clear which confounders were adjusted for and why they were included

(b) Report category boundaries when continuous variables were
categorized

(c) If relevant, consider translating estimates of relative risk into
absolute risk for a meaningful time period

Other analyses 
--------------

Report other analyses done—eg analyses of subgroups and interactions,
and sensitivity analyses 
-->

You can include code in this document like this:

```{r main, echo=FALSE}
source("main.R") ## This "imports" the main script file of your project and run any code in it
```

You can also embed plots:

```{r plot, echo=FALSE}
plot(pressure)
```

You can also mix text and code, so called inline code, like this:
`r 2+5`.

# Discussion

The aim of this study was to evaluate the performance of individual AFs in predicting OFIs and 
to assess their suitability as a screening tool for MMCs. This study showed that no AFs was significantly successful in predicting OFIs. Four AFs ("SBP<90," "dead at 30 days," "GCS<9 and not intubated," and "CPR and thoracotomy") were equivalent to random chance in their predictive ability. The highest AUC was for ">60 minutes until first intervention,". Notably, all AFs performed significantly below the accepted standard for diagnostic tests (>0.8).

## Limitations

OFI, while dichotomously defined, includes a diverse set of outcomes.
This convolutes the process of creating AFs that perform well in
predicting OFIs. AFs would favour the identification of some, but not
all, errors.

Each AF was individually related to OFI. Consequently, if an AF
demonstrated a true positive correlation with an OFI, instances where
the same patient case had multiple AFs introduce uncertainty regarding
which specific AF truthfully predicted the OFI.

The number of missing values varied greatly among the AFs. As stated
above, missing values from each AF were coded as "not flagged" by an AF.
This means that some instances where patient cases were not flagged,
might actually meet the AF criteria, but for some reason not documented.
This reduces the validity of some AFs that featured high amounts of
missing data.

##Interpretation
Higher ISS, longer times to CT, longer times to first intervention, and higher levels of care correlated with OFIs, consistent with a recent study by Albaaj et al. [@albaaj2023patient]. Notably, GCS at the ED also showed a correlation with OFIs despite identical median GCS values, likely due to distribution differences outside the IQR.

The AF "GCS\<9 and not intubated" showed no significant ability to
predict OFIs, similarly shown by Willis et al. [@willis2008interpreting]. Possibly because the AF might be more reflective of clinical
decision rather than quality of care [@willis2008interpreting]. For instance, 
An abnormal GCS resulting from intoxication is managed differently than one caused by a head injury.

Both delay-related AFs: "\>60 until first intervention" and "\>30min
until first CT", showed a correlation with OFI. Comparably, Teixeira et al. [@teixeira2007preventable] linked delays
to preventable death and Ghorbani et al.
[@ghorbani2018analysis] identified treatment and CT delays as common errors in trauma care. A recent study in Japan by Yamamoto
et. al [@yamamoto2023immediate] found reduced mortality with whole-body CT scans within 10 minutes of ED arrival.

"Dead at 30 days"
and "CPR and thoracotomy" showed no correlation with OFI. This could be
explained by inevitable mortality, thus presenting with a smaller time
frame for mistakes to be made.

AFs
generally exhibited a high specificity relative to their sensitivity
which may be attributed to the outcome being rather uncommon, whilst
every individual AF misses the majority of OFIs. This could suggest that
the AFs are too rigid in their criteria. On the contrary, a too
inclusive criteria might flag several patients, without an OFI [@willis2008interpreting].

## Generalizability

AFs are widely used, but lack consensus on which to use, purpose, and application [@stelfox2010quality; @stelfox2011evidence;
@berg2022perceived]. They have static criteria and may not adapt to evolving healthcare challenges, or upon resolution of the quality gaps they were designed for, reducing their utility over time. This underscores the need for regular evidence-based reviews of AF performance [@santana2014development]. However, these reviews might be cost-prohibitive

While AFs can help identify an OFI, they might overshadow other contributing factors, leading to over-reliance and potentially failing to recognize cases outside their scope.
[@willis2008interpreting]. For example, AFs often focus on aspects relevant to doctors, despite the interdisciplinary nature of trauma care. Lastly, Confounding variables, such as different treatment times based on injury severity, add to the uncertainty of AFs' effectiveness.
[@willis2008interpreting].

Due to the
study being conducted in a single, trauma level I centre in Stockholm it
is unrealistic to assume applicability of these findings at other
hospitals facing unique challenges [@world2009guidelines]. For instance AFs that benefit
low and middle income may not benefit
high income countries and vice versa [@berg2022perceived].

## Future Studies

Given the expanding field of artificial intelligence and machine
learning, new studies should focus on using these tools to screen for
patient cases at risk of an OFI. This will reduce misclassification
and maintain high performance, attributable
to an algorithm's capability to adapt. A recent preprint by Attergrim et al.
[@attergrim2023predicting] found that machine learning
models outperformed the AFs currently used at Karolinska University
Hospital in predicting OFIs. Should the use of AFs remain, they must be routinely updated to reflect current
quality gaps. 
<!--

Key results
-----------
Summarise key results with reference to study objectives

Limitations
-----------

Discuss limitations of the study, taking into account sources of
potential bias or imprecision.  Discuss both direction and magnitude
of any potential bias

Interpretation
--------------

Give a cautious overall interpretation of results considering
objectives, limitations, multiplicity of analyses, results from
similar studies, and other relevant evidence

Generalisability
----------------

Discuss the generalisability (external validity) of the study results

-->

# Conclusion

AFs are widely used as a tool in QI programmes and
should reflect the current gap in trauma care quality. Strong evidence
supporting their efficacy in quality improvement is scarce. In this
study, no audit filter seemed to show promising performance in
predicting OFIs with AUC ranging from 0.5 to 0.63. These results
highlight the importance of continually evaluating the relevance of AFs.
However, this process takes time and resources and using an alternative
to AFs is recommended.

# References

Keep? However,
according to Sanddal et al. [@sanddal2011analysis] airway management was
associated with OFI, suggesting that an alternative AF related to airway
management might perform differently.

Notably, OFI in this study was not soley a reflection
of mortality but also morbidity. Rather, it could highlight differences in quality gaps
between hospitals.