### 2024-09-14 manuscript.Rmd
update

---
title: Performance of Individual Audit Filters in Predicting Opportunities for Improvement
  in Adult Trauma Patients
author: "Josef Al-Khalili"
output:
  html_document: default
  word_document: default
  pdf_document: default
subtitle: "A single-center registry-based retrospective cohort"
bibliography: bibliography.bib
csl: vancouver.csl
editor_options:
  markdown:
    wrap: 72
---

```{r sourcecode, echo=FALSE, fig.align="center", warning=FALSE, message=FALSE,error=FALSE}
source("main.R")
library(rofi)
library(dplyr)
library(ggplot2)
library(pROC)
library(gtsummary)
library(gt)
library(boot)
library(knitr)
library(webshot2)
```

# Abstract

## Background

Trauma audit filters, based on standards of care, serve as screening
tools for patient cases to be assessed at interdisciplinary mortality
and morbidity conferences. These conferences aim to identify potential
opportunities for improvement in the care of trauma patients. This study
aimed to determine the effectiveness of individual audit filters in
predicting opportunities for improvement and their relevance as
screening tools for morbidity and mortality conferences.

## Methods

We conducted a single-centre registry-based retrospective cohort study based on register data collected between 2013 and 2022 from the Karolinska University Hospital in Sweden. Performance was measured using sensitivity, specificity and Cohen’s kappa coefficient for ten audit filters. Bootstrapping with 1,000 iterations calculated 95% confidence intervals.

## Results

Our analysis included 8,309 patients. Audit filters showed a Cohen’s kappa coefficient ranging from -0.07 (95% CI [-0.14-0]) for “GCS < 9 and not intubated” and 0.08 (95% CI [0.05-0.12]) for “Massive transfusion”. The Cohen’s kappa coefficient for all audit filters combined was 0.04 (95% CI[0.03-0.04]).

## Conclusion

No individual audit filter performed well in predicting opportunities for improvement, which suggests reassessing the current method of screening for mortality and morbidity conferences.

# Introduction

Trauma, defined as a physical injury and the body's associated response,
causes 8-9% of global deaths annually [@world2014injuries;
@roth2018global] and ranks highest in disability-adjusted life-years
(DALYs) between ages 10-49 years [@vos2020global], impacting
individuals, communities and society at large [@newnam2014impacts; @david2022measuring]. Road
traffic injuries alone cost approximately 2% of gross domestic product
(GDP) in high-income countries [@world2014injuries]. Despite
advancements in trauma care, preventable deaths [@kwon2014pooled;
@ghorbani2018analysis; @teixeira2007preventable; @vioque2014classifying] and errors persist [@nikouline2021errors;
@vioque2014classifying; @o2013opportunities; @teixeira2007preventable]. To address this, trauma quality
improvement programs [@American2022resources], commonly including
mortality and morbidity conferences [@world2009guidelines] have been
implemented. The conferences have shown to foster error reduction and
encourage reporting [@kong2016analysis] without negative feedback
[@lazzara2022morbidity].

Audit filters are criteria used to identify cases for mortality
and morbidity conferences aiming to pinpoint potential opportunities for
improvement [@American2022resources]. Audit filters can involve outcomes
such as "death" or adherence to guidelines, such as "delay to treatment"
[@world2009guidelines]. Opportunities for improvement are typically
associated with failures in initial care [@vioque2014classifying] like
airway management, fluid resuscitation, haemorrhage control and chest
injury management [@sanddal2011analysis; @roy2017learning;
@o2013opportunities]. There seems to be a lack of common, evidence-based
set of audit filters for evaluating trauma care quality
[@stelfox2010quality], and while early studies reported mortality
reduction [@chadbunchachai2001comparative; @chadbunchachai2003study],
newer studies differ [@evans2009audit; @willis2008interpreting]. Moreover, the use of audit filters has been
associated with high frequencies of false positives, ranging from 24% to
80% in different contexts [@sanddal2011analysis; @roy2017learning; @ghorbani2018analysis]. Few
studies have focused on
non-mortality outcome measures, for evaluating performance of audit
filters but the evidence remains inconsistent [@di2008acscot;
@glance2012association] and
some audit filters do not appear to correlate with opportunities for
improvement at all [@lewis2017delay]. However, the perceived
usefulness of audit filters remains high [@berg2022perceived]. The aim of this study is to determine the performance of individual
audit filters in predicting
opportunities for improvement and their relevance as a
screening tool for mortality and morbidity conferences

# Methods

## Study design

We conducted a single-centre registry-based retrospective cohort study
of all trauma patients included in both the Karolinska University
Hospital trauma registry and trauma care quality database between 2013
and 2022 to compare the performance of audit filters in their ability to
predict opportunities for improvement. This study was approved by the
Swedish Ethical Review Authority, approval number 2021-02541 and
2021-03531.

## Setting

Karolinska University Hospital manages around 1,500 trauma patients
annually. The Karolinska Trauma Registry, which reports to the national
trauma registry SweTrau [@SweTrau], includes all patients that result in
trauma team activation or who have an Injury Severity Score (ISS) of
more than 9. This registry includes data on vital signs, timings,
injuries, interventions, and patient demographics as per the European
consensus statement [@ringdal2008utstein]

The trauma centre at the Karolinska University Hospital also keeps a
local care quality database with information intended for the mortality
and morbidity conferences, including identified triggered audit filters
and opportunities for improvement. Opportunities for improvement are
assessed in multiple stages of scrutiny. Notably, mortality is directly
referred to the mortality and morbidity conferences to assess
opportunities for improvement.

During the study period, the screening of patient cases for potential
opportunities for improvement was formalised. Prior to 2017, not all
patient cases were subjected to an initial evaluation by a specialised
nurse. However, following 2017 all patient cases underwent screening.
Importantly, audit filters and an initial evaluation by a specialised
nurse has consistently functioned as the screening tool for mortality
and morbidity conferences to identify potential opportunities for
improvement.

## Participants

The study included all patients treated at Karolinska University
Hospital from 2013 to 2022 who underwent screening for opportunities for
improvement. Patients aged under 15 years were excluded due to differences in
clinical management.

## Variables and data sources/measurements

The outcome was opportunities for improvement, as established by the
mortality and morbidity conference, and was treated dichotomously as
"Yes - At least one opportunity for improvement identified" and "No - No
opportunity for improvement identified".

The ten audit filters used at the Karolinska University Hospital served
as the exposure variables. For each audit filter, we identified the cohort of patients in which the filter was applicable, and evaluated the filters’ performance in that cohort: For “ISS > 15 and no team activation” and “ISS > 15 and not in ICU” we evaluated the filters’ performance in patients with an ISS greater than 15; for "GCS < 9 and not intubated" we analyzed patients with a GCS less than 9; and lastly, for "No anticoagulants within 72 hours after TBI" we identified patients with head trauma as defined by the precence of at least one head injury according to available AIS codes.

The data was retrieved from the Karolinska Trauma Registry and trauma
care quality database. Patient data on vital signs, care processes and
interventions, level of care and time aspects was retrieved from the trauma registry whereas the exposure, audit filters and the outcome, and opportunities
for improvement were retrieved from the quality database.

## Statistical methods

We used R [@R2022language] for statistical analysis. Patient
characteristics for cases with and without opportunities for improvement
were compared using the Wilcoxon signed-rank test for continuous
variables and the Chi-squared test for categorical variables. The
predictive performance of audit filters was assessed by calculating the
sensitivity, specificity and Cohen's kappa for each audit filter in
predicting opportunities for improvement. Bootstrapping with 1,000
iterations was applied to estimate 95% confidence intervals for all
performance metrics. We selected cohorts relevant to each filter: For
"ISS \> 15 and no team activation" and "ISS \> 15 and not in ICU" we
examined patients with an ISS greater than 15. For "GCS \< 9 and not
intubated," we focused on patients with a GCS less than 9. Lastly, for
"No anticoagulants within 72 hours after TBI" we identified patients
with head trauma as defined by the presence of at least one head injury
according to available AIS codes. Missing values were numbered but
ignored in the analysis.

# Results

After excluding patients aged less than 15 years and patients not screened
for opportunities for improvement, a total of 8,309 individuals were
included in the study. Out of these 8,309 patients, a total of 512 (6.2%) patients had an
opportunity for improvement, see Figure 1.

Patients with an opportunity for improvement had significantly higher
median age,
`r inline_text(tableOne, variable = pt_age_yrs, column = stat_2)` years vs
`r inline_text(tableOne, variable = pt_age_yrs, column = stat_1)` years,
`r inline_text(tableOne, variable = pt_age_yrs, column = "p.value")`, median 
ISS `r inline_text(tableOne, variable = ISS, column = stat_2)` vs
`r inline_text(tableOne, variable = ISS, column = stat_1)`,
`r inline_text(tableOne, variable = ISS, column = "p.value")` and
intubations in the emergency department (ED), 
`r inline_text(tableOne, variable = ed_intubated, level = "Yes", column = stat_2)` patients
vs
`r inline_text(tableOne, variable = ed_intubated, level = "Yes", column = stat_1)` patients,
`r inline_text(tableOne, variable = ed_intubated, column = "p.value")`.
Patients with opportunity for improvement had longer times to first major intervention from hospital arrival median
`r inline_text(tableOne, variable = dt_ed_emerg_proc, column = stat_2)`
minutes vs
`r inline_text(tableOne, variable = dt_ed_emerg_proc, column = stat_1)` minutes,
`r inline_text(tableOne, variable = dt_ed_emerg_proc, column = "p.value")`.
and longer times to first CT, median
`r inline_text(tableOne, variable = dt_ed_first_ct, column = stat_2)` minutes vs
`r inline_text(tableOne, variable = dt_ed_first_ct, column = stat_1)` minutes,
`r inline_text(tableOne, variable = dt_ed_first_ct, column = "p.value")` minutes.
The frequency of opportunities for improvement was highest in patients
admitted to the intensive care unit,
`r inline_text(tableOne, variable = host_care_level, level = "Intensive care unit", column = stat_2)`
patients, compared to general ward
`r inline_text(tableOne, variable = host_care_level, level = "General ward", column = stat_2)`. Notably, there was a statistically significant
correlation between GCS at ED and opportunities for improvement (p \<
0.001).
Gender, mean systolic blood pressure, death at 30 days and trauma team
activation were not significantly correlated with opportunities for
improvement (p\>0.05). The highest amount of missing data was observed for the audit
filter "time to first major intervention", which was missing for 5,990 patients, (72.1%). See Table 2
for details of selected patient characteristics.

## Performance of Individual Audit Filters

The number of patient cases flagged by each individual audit filter
varied between (n = 107, 27.4%) for "GCS \< 9 and not intubated" and (n
= 3907, 53.5%) for "\>30 min until CT". The highest number of missing
values was observed with "\> 60 min until first intervention" (n = 5990,
72,1%). The lowest number of missing values was observed with "dead at
30 days" and "ISS \> 15 and not in ICU" (n = 11, 0.1%). See table 1 for
missing values for each audit filter. The largest cohort was observed
with "Dead at 30 days" (n = 8298, 99.9%) and the smallest cohort was
observed with "GCS \< 9 and not intubated" (n = 390, 4.7%). In total
5977 out of 8309 (71.9%) patient cases were flagged by at least one
audit filter.

The audit filter with the highest sensitivity was "\> 60 min until first
intervention" (83.3%, 95% CI [79.2-87.5]). The lowest sensitivity was
seen in "SBP \< 90" (8%, 95% CI [5.6-10.2]) and "Dead at 30 days" (8%,
95% CI [5.6-10.2]). The audit filter with the highest specificity was
"Massive transfusion" (95.8%, 95% CI [95.4-96.3]). The lowest
specificity was seen in "ISS \> 15 and no team activation" (21.6%, 95%
CI [19.9-23.2]). The sensitivity for all audit filters combined was
92.6% (95% CI [90.5-95]) and the specificity was 29.4% (95% CI
[28.4-30.5]).

The highest Cohen's kappa was seen for the audit filter "Massive
transfusion" (0.08, 95% CI [0.05-0.12]). The smallest Cohen's kappa was
seen in "GCS \< 9 and not intubated" (-0.07, 95% CI [-0.14-0]. The
Cohen´s kappa for all audit filters combined was 0.04 (95% CI
[0.03-0.04]. See table 3 for the performance of all audit filters using
sensitivity, specificity and Cohen's kappa coefficient.

```{r fig1, echo=FALSE, fig.align="center", warning=FALSE, message=FALSE, error=FALSE}
knitr::include_graphics("ofi-flowchart.pdf")
```

# Discussion

This study  shows that audit filters, individually or in combination,  were not effective in predicting opportunities for improvement. As a result, the development of a new screening method to better identify potential opportunities for improvement is advised.

agree

disagree
Some previous studies have shown similar findings to this study. Lewis et al. looked at one audit filter, "delay to operating room" of more than 2 hours and its correlation with opportunities for improvement. They found that the audit filter had no significant ability to predict opportunities for improvement. Similarly, the delay-related audit filters: "\>60 min until first intervention" and
"\>30min until first CT" showed noability to predict opportunities for improvement. A big meta analysis by Evans et al. sought to identify a common set of evidence based audit filters but no study fulfilled the inclusion criteria.
A 2008 study by Di Bartolomeo et al. [@di2008acscot] looked at 4 ASCOT audit filters in relation to 30-day mortality and a morbidity measure. One audit filter showed a correlation with mortality but no audit filter showed a correlation with the morbidity measure.

A recent study by Berg et al. [@berg2024effects], looking at implementation of audit filters on mortality concluded that audit filters might be useful in reducing the mortality. 
Previous studies suggest poor performance of audit filters for improving
trauma care [@evans2009audit; @stelfox2010quality; @willis2008interpreting; @di2008acscot; @glance2012association; @lewis2017delay]. This study adds to that evidence by showing that audit filters perform poorly in predicting opportunities for
improvement encompassing mortality and non-mortality
associated outcomes. Some audit filters might be more reflective of clinical decisions rather than quality of care [@glance2012association]. For instance, an abnormal GCS because of intoxication
is managed differently to an abnormal GCS due to head injury, which could explain the poor performance of "GCS < 9 and not intubated", similarly shown by
Willis et al. [@willis2008interpreting]. This could also explain the poor performance of the delay-related audit filters: "\>60 min until first intervention" and
"\>30min until first CT". Similarly, Lewis et al. [@lewis2017delay] showed that delay
to the operating room was not associated with opportunities for improvement.
In contrast, Teixeira et al. [@teixeira2007preventable] identified delays to
treatment and delays to CT as common errors in trauma care. A recent
study in Japan by Yamamoto et. al [@yamamoto2023immediate] observed
significantly reduced mortality with whole-body CT scans within 10
minutes of arrival to the ED. If anything, this highlights the unique opportunities for improvement in different hospitals. The audit filters also varied greatly in terms of expected mortality. Some audit
filters such as "Massive transfusion" and "Dead at 30 days" have a high
expected mortality. In such cases the reduced time frame for any potential
opportunity for improvement to occur could explain their poor performance. When audit filters were combined they had a low specificity relative to its sensitivity which corresponds to previous studies showing high numbers of false positives [@sanddal2011analysis; @roy2017learning; @ghorbani2018analysis]. The sensitivity and specificity for individual audit filters varied substantially, suggesting that some audit filters are too rigid, missing the majority of opportunities for improvement while some are too inclusive, causing high numbers of false positives [@willis2008interpreting].

There were some limitations in this study. Firstly, it is unlikely that the specific associations presented in this study are generalizable to other hospitals with similar systems, but the broader finding that audit filters are poorly associated with opportunities for improvement is likely to be externally valid [@world2009guidelines]. Secondly, instances where the
same patient case had multiple audit filters introduce uncertainty
regarding which specific audit filter truthfully predicted an
opportunity for improvement. Finally, the number of missing values
differed significantly among the audit filters, resulting in varying
degrees of validity for each filter.

The static nature of audit filters inherently limit their adaptability to
emerging healthcare challenges and resolved quality gaps, reducing their
utility over time. Therefore, regular evidence-based reviews are needed
to ensure performance [@santana2014development], though such reviews may
be cost-prohibitive. Furthermore, they might overshadow other
contributing factors to opportunities for improvement
[@willis2008interpreting].

In summary, audit filters are widely used as a tool in quality
improvement programmes and should reflect current gaps in trauma care
quality. In this study, no individual audit filter or their combination were
successful in predicting opportunities for improvement. Given these
results, future studies should focus on developing new tools with higher predictive
ability.

# Acknowledgements

The authors thank all professionals taking part in the monthly mortality
and morbidity conferences.

# References